Modelling potential environmental impacts of science activity in Antarctica

### 2025-09-24 talk/index.qmd
final version of talk - made html non-self-contained

---
title: "Modelling potential environmental impacts of science activity in Antarctica"
title-slide-attributes:
    data-background-image: background-image.png
    data-background-opacity: "0.65"
author: 
  - name: David O'Sullivan
    affiliations: 
      - name: Geospatial Stuff
      - name: University of Auckland
  - name: Fraser Morgan
    affiliations: 
      - name: Bioeconomy Science Institute
format:
  revealjs:
    theme: [black, custom.scss]
    transition: none
    navigation-mode: vertical
    fragment-in-url: true
    slide-number: false
    logo: logo.png
    footer: "Palmerston North Regional GIS Forum, September 2025"
    self-contained: true
---

@@

## Up $\neq$ down {background-image="klondike.jpg" background-opacity="0.65"}
### Slope is critical to movement, so attach heights to nodes and calculate slope in each direction along lattice edges
### Then apply a *hiking function* to estimate traversal times

@@

## Completing and applying the network
### Assign edges *estimated traversal times* based on slope, the hiking function, and land cover (moraine or rock)
### Make the nodes and edges into a _directed graph_
### Then use various _graph algorithms_ to find, e.g., everywhere-to-everywhere shortest paths

@@

##
![](fig-hiking-network-example-1.png){fig-align="center"}

## Betweenness centrality
### After some exploration, we landed on _betweenness centrality_ as a useful metric
### Counts how often each node appears on shortest paths between every other pair of nodes
&nbsp;

### &rarr; Indicator of relative likelihood of locations being visited

@@

---

## [Conclusions]{style="text-shadow:0px 0px 8px rgba(102,0,0,0.5);"} {background-image="beacon-valley.jpg" style="text-shadow:0px 0px 8px rgba(102,0,0,0.5);"}
### Terrain-differentiated data-fitted hiking functions are a novelty
### Potential wider application of radius-limited betweenness centrality?
### Manuscript in preparation

::: aside
Image: Beacon Valley, from [reddit.com/rHumanForScale](https://www.reddit.com/r/HumanForScale/comments/by9q7f/beacon_valley_antarctica/)
:::

@@

::: {style="background-color:rgba(102,0,0,0.1);"}
<h2 style="text-align:center;text-shadow:0px 0px 8px rgba(102,0,0,0.35),0px 0px 4px rgba(102,0,0,0.65);">For more</h2>
### On this project [dosull.github.io/antarctica](https://dosull.github.io/antarctica) {style="text-align:center;text-shadow:0px 0px 8px rgba(102,0,0,0.35),0px 0px 4px rgba(102,0,0,0.65);"}
### About ![](logo.png){style="vertical-align:middle;width:60px;text-align:center;"} Geospatial Stuff [dosull.github.io](https://dosull.github.io) {style="text-align:center;text-shadow:0px 0px 8px rgba(102,0,0,0.35),0px 0px 4px rgba(102,0,0,0.65);"}
:::

>>>>

---
title: "Modelling potential environmental impacts of science activity in Antarctica"
title-slide-attributes:
    data-background-image: background-image.png
    data-background-opacity: "0.65"
author: 
  - name: David O'Sullivan
    affiliations: 
      - name: Geospatial Stuff
      - name: University of Auckland
  - name: Fraser Morgan
    affiliations: 
      - name: Bioeconomy Science Institute
format:
  revealjs:
    theme: [black, custom.scss]
    transition: none
    navigation-mode: vertical
    fragment-in-url: true
    slide-number: false
    logo: logo.png
    footer: "Palmerston North Regional GIS Forum, September 2025"
    self-contained: false
---

@@

## Up $\neq$ down {background-image="klondike.jpg" background-opacity="0.65"}
### Slope is critical to movement, so attach heights to nodes and calculate slope in each direction along lattice edges
### Then use a *hiking function* to estimate traversal times

@@

## Completing and applying the network
### Estimate edge *traversal times* based on slope, hiking function, and land cover (moraine or rock)
### Make nodes and edges into a _directed graph_
### Use various _graph algorithms_ to find, e.g., everywhere-to-everywhere shortest paths

@@

##
![](hiking-network-example.png){fig-align="center"}

##
![](single-source-shortest-paths.png){fig-align="center"}

##
![](single-source-shortest-paths-2.png){fig-align="center"}

## Betweenness centrality
### After some exploration, we landed on _betweenness centrality_ as a useful metric
### Counts how often each node appears on shortest paths between every other pair of nodes {.fragment}
&nbsp;

### &rarr; Indicator of relative likelihood of locations being visited {.fragment}

@@

# Conclusions {.smaller background-image="beacon-valley.jpg"}

[Image: Beacon Valley, from [reddit.com/rHumanForScale](https://www.reddit.com/r/HumanForScale/comments/by9q7f/beacon_valley_antarctica/)]{.aside .position bottom=0 left=0}

## {background-image="beacon-valley.jpg"}

::: {style="text-shadow:0px 0px 8px rgba(102,0,0,0.35),0px 0px 4px rgba(102,0,0,0.65);background-color:rgba(102,0,0,0.1);backdrop-filter:blur(2px);padding:5px;border-radius:5px;"}
### Hiking network concept is implicit in hiking functions but a new terrain representation
### Terrain-differentiated GPS data-fitted hiking functions are novel {.fragment}
### Potential wider application of radius-limited betweenness centrality {.fragment}
### Next up: start a conversation with Antarctic scientists about planning {.fragment}
:::

@@

::: {style="background-color:rgba(102,0,0,0.1);backdrop-filter:blur(2px);padding:5px;border-radius:5px;"}
<h2 style="text-align:center;text-shadow:0px 0px 8px rgba(102,0,0,0.35),0px 0px 4px rgba(102,0,0,0.65);">For more</h2>
### On this project [dosull.github.io/antarctica](https://dosull.github.io/antarctica) {style="text-align:center;text-shadow:0px 0px 8px rgba(102,0,0,0.35),0px 0px 4px rgba(102,0,0,0.65);"}
### On ![](logo.png){style="vertical-align:middle;width:60px;text-align:center;"} Geospatial Stuff [dosull.github.io](https://dosull.github.io) {style="text-align:center;text-shadow:0px 0px 8px rgba(102,0,0,0.35),0px 0px 4px rgba(102,0,0,0.65);"}
:::
